In the beginning of this thesis, we introduced the Boltzmann equation and the kinetic models approximating the Boltzmann equation namely, the BGK and ES-BGK equations. We then reduced the kinetic equations to one dimension in spatial variable by assuming symmetry in the other two dimensions.

We applied the DG discretization in velocity to the reduced kinetic model equations and the Boltzmann equation. First and second order FV methods were developed in the spatial variable of the kinetic models. Splitting methods were used to solve the homogeneous and particular parts of the kinetic models separately. The fractional step method was used to split the first order method and Strang splitting was used for splitting the second order method.

The CLAWPACK software package was extended for use with the BGK and ES-BGK kinetic equations of the Boltzmann equation. We took advantage of the splitting techniques used by CLAWPACK and provided a first and second order Riemann solver for the software. In addition, we supplied components that perform discretization in velocity.

Using the CLAWPACK software, we then verified the order of convergence of the first and second order FV methods by solving the homogeneous transport equation with these methods and compared the results to the exact solution. Numerical results of the heat transfer and shock wave problems were also carried out. The heat transfer problem consisted of bringing the state of a gas in a closed system to equilibrium. We verified that our methods conserved mass. The shock wave problem consisted of an open system with a constant incoming stream and outgoing stream of a gas modeled by Maxwellian distributions with parameters determined by the Rankine-Hugoniot relations \cite{roshko}. Within the domain, a shock wave is formed. The density and temperature profiles as well as the reciprocal thickness of the shock wave solutions were compared to Alsmeyer's \cite{alsmeyer} experimental results and the order of convergence was verified for the first and second order solvers. For future work, CLAWPACK can be extended for solving these types of problems in two dimensions.

A velocity dependent collision frequency BGK model was developed based on \cite{tch}. The BGK model approximation was developed with the assumption that the state of the gas was near a Maxwellian. The new model was designed to correct for distributions that are too far from the Maxwellian to be approximated by the BGK or ES-BGK models. With this new model, we were able to attain the same relaxation rate of the directional temperature as the Boltzmann equation. For future work, the relaxation rates of higher moments of the Boltzmann equation can be enforced. In addition, the extension to include the spatial domain can be considered for problems including solutions to the shock wave problem. This can assist in producing accurate models of the shock wave problem where the distribution of the gas at the shock front is far from a Maxwellian.
